Thanks to the effortless sharing of video files, there is a relatively big community of movie and TV series fans on the Internet who spend a significant amount of their free time translating the subtitles of foreign movies to their native languages.


There are many sophisticated commercial tools available that assist professional translators with their translation tasks, as well as software solutions used by big translation agencies; however, a complex tool for fan-based translation using  state-of-the-art techniques is still missing.

In contrast to the area of commercial translation, where each translation agency keeps previously translated texts as proprietary knowledge, fan translations are based on a non-commercial community approach that creates huge amounts of translated subtitles that are publicly available. There has not been a tool for subtitle translation taking advantage of the public availability of this data.

The subtitles themselves are an ideal source of parallel data for linguistic research. They are used for the creation of  parallel corpora for statistical machine translation (e.g.\ the CzEng corpus\footnote{\url{http://ufal.mff.cuni.cz/czeng/}} used for training the statistical machine translation tools at ÚFAL). The film industry is also aware of this fact and there have been some attempts to do an automatic translation of  subtitles  with minimal post-editing based on existing subtitles data.\footnote{Marian Flanagan (2009): \emph{Using Example-Based Machine Translation to translate DVD Subtitles.} Proceedings of the 3rd Workshop on Example Based Machine Translation, p. 85–92}

Translation memories (discussed in more detail in Section~\ref{sec:translation_memories}) are in these days the most common tool used by translators. The memories are usually very domain specific to make the computation feasible on the translators' PC and to avoid any confusion in terminology. In the case of movie subtitles, there is no such terminological danger. Moreover, we believe that the similarity of the movie scripts will produce more data, leading to better results. These arguments make us believe that creating a translation memory based tool focusing on movie subtitle translation can be both useful and successful among the target users.

As indicated in the previous paragraphs, the goal of our project is to create an application that will help amateur translators of movie and TV show subtitles with their efforts. The core of the application is a large translation memory, which will be gradually extended and improved by users' translations. The translation memory exists only in one publicly available instance, which is on the server, and is therefore shared by all users. We focus on the English-Czech language pair; the application itself is language independent,
% I dont understand this: However, training data for statistical natural language processing tools are needed
but different training data would be necessary
to run the application for another language pair.

Although our original goal was only to build a translation memory, we found that the collected data can be used in building a parallel corpus for training a statistical machine translation system. We built such a system using the Moses machine translation engine and added it into our project as another option for translators; we describe the system in Chapter~\ref{chap:moses}.

The collected data can be also used for research on the language used in the subtitles in general.

The following chapter contains a general introduction to the area of producing subtitles and of computer-assisted human translation. The next chapter provides general information about implementing the project, including usage of external libraries. The following chapters contain detailed description of the individual project modules.

The second part of the documentation is devoted to the evaluation of the development process of the project, including the reasons for the decisions we took, and an overview of the work distribution among the members of the project.

The last part of the documentation contains the manuals. Chapter~\ref{chap:technical_manual} contains a manual for a server administrator, explaining how to install the server part of the application. Chapter~\ref{chap:users_manual} contains the manual for web application users (which is also encorporated into the application as a Help page).

We would like to thank doc. RNDr. Vladislav Kuboň, Ph.D. for useful and practical
advices during supervising the FilmTit project, Mgr. Milan Fučik for technical
support of our virtual machine, RNDr. Ondřej Bojar, Ph.D. for advice with Moses,
the admin of opensubtitles.org for the data, the testers for finding many interesting bugs,
and KFC and Burger King for free wifi connection.
